---
title: "Chapada Checks"
author: "TE Technicians"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: espresso
---

```{=html}
<style>
.vscroll-plot {
    width: 1000px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
body {
  color: dark brown;
}
</style>
```

```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
library(ggpubr)
library(gghighlight)
library(kableExtra)
library(ggrepel)
library(knitr)
library(utils)
library(pacman)

```

```{r set knitr options,include = F}
options(width = 150)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height, options$max.width)
    )
    hook_output(x, options)
  })
})
```

> This RMD was created to help staff keep tabs on the technical operation of the experiement, if you'd like to explore the data, check out the project dashboard:

```{=html}
<a href="https://aslewis.shinyapps.io/chapada_dashboard/" target="_blank">
  <button class="btn btn-primary">Chapada Dashboard!</button>
</a>
```

```{r load data, echo = F}
#read in files from dropbox
setwd("C:/Users/LaGorgaL.S/Documents/Chapada_Stem/daily_checks/")
dbfold <- "C:/Campbellsci/LoggerNet/CHAPADA_Loggerfiles/"
files <- list.files(dbfold, pattern = "*.dat", full.names = TRUE)
names(files) <- gsub(".dat", "", basename(files))

#read into list to stay organized, format head function
chapada_data <- lapply(files, function(file) {
 data <- fread(file,skip=1, header = TRUE)[-c(1,2)]
  return(data)
})
names(chapada_data) <- names(files)

#convert timestamps
extended_data <- lapply(chapada_data, function(df) {
  df$TIMESTAMP <- as.POSIXct(df$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S")
    df<- df %>% filter(TIMESTAMP>=format(floor_date(Sys.time(),unit="month"),"%Y-%m-%d %H:%M:%S"))
  return(df)
})

chapada_data <- lapply(chapada_data, function(df) {
  df$TIMESTAMP <- as.POSIXct(df$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S")
  df<- df %>% filter(TIMESTAMP>=format(Sys.time()-259200,"%Y-%m-%d %H:%M:%S"))
  return(df)
})


```
<br>
<br>

### Battery Voltages for this month! 12-14V is ideal... If you notice that the battery voltage is drifting, it may be time to change the battery.

```{r batteries babyyyy, echo=F}
#aggregate and make long battery table for plotting
battery_tables <- grep("CheckTable|ChamberBME280|Redox15|Teros12Table", names(extended_data), value = TRUE)
batt_plot_data <- bind_rows(lapply(battery_tables, function(table) {
  df <- extended_data[[table]]
  df %>% select(TIMESTAMP, BattV = matches("BattV|BattV_Min|BattV_Avg")) %>%
    mutate(origin = table)
}))

ggplot(batt_plot_data, aes(x = TIMESTAMP, y = as.numeric(BattV), color = origin))+
  geom_line(na.rm = TRUE) +
  labs(title = "Battery Voltage Across Tables", y = "Voltage (V)", x = "Time") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) 
```
<br>
<br>

### The plot below shows this months' record of the ringdown times for both licors. Make sure that you are not seeing any long-term drift.

```{r ringydingy!, echo=F}
ringdown <- grep("FLUX_7810", names(extended_data), value = TRUE) 
ringdown_data <- bind_rows(lapply(ringdown, function(table){
  df <- extended_data[[table]]
  df %>% select(TIMESTAMP, ringdown = matches("ringdown")) %>%
    mutate(origin = table)
  }))

ggplot(ringdown_data, aes(x = TIMESTAMP, y = as.numeric(ringdown), color = origin)) +
  geom_line(na.rm = TRUE) +
  labs(title = "Ringdown Time by Instrument", y = "Ringdown", x = "Time") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) 
```
<br>
<br>

### Below are the CO2 and CH4 fluxes for the past 3 days
```{r flux deluxe!, echo=F}
flux_comb_tables <- grep("FLUX_COMB", names(chapada_data), value = TRUE)

flux_data <- bind_rows(lapply(flux_comb_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, Fluxing_Chamber, Flux_Status, Manifold_Timer, H2O, CO2, CH4) %>%
    mutate(origin = table)
}))

ggplot(flux_data, aes(x = TIMESTAMP, y = as.numeric(CO2), color = as.factor(Fluxing_Chamber))) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "CO2 Concentration by Chamber", y = "CO2 ppm", x = "Time") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) 

ggplot(flux_data, aes(x = TIMESTAMP, y = as.numeric(CH4), color = as.factor(Fluxing_Chamber))) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "CH4 Concentration by Chamber", y = "CH4 ppm", x = "Time") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) 
```
<br>
<br>

### Flux Slopes and R2 Values of plotted 

```{r flux flag 2 - high, echo=F}
#flux slopes in table form
high <- chapada_data[["Chapada_High_FLUX_COMB"]]
high <- high %>% filter(Flux_Status == 2, Manifold_Timer >= 180 & Manifold_Timer <= 540) 
high_day <- high %>% filter(hour(TIMESTAMP)>= 6 & hour(TIMESTAMP)<= 18)
high_night <- high %>% filter(hour(TIMESTAMP)<= 6 | hour(TIMESTAMP)>= 18)

lm_high_day <- high_day %>%
  group_by(Fluxing_Chamber) %>%
  summarise(
    Day_CO2_slope = coef(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))[2],
    Day_CO2_r2 = summary(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))$r.squared,
    Day_CH4_slope = coef(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))[2],
    Day_CH4_r2 = summary(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))$r.squared
  )

lm_high_night <- high_night %>%
  group_by(Fluxing_Chamber) %>%
  summarise(
    Night_CO2_slope = coef(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))[2],
    Night_CO2_r2 = summary(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))$r.squared,
    Night_CH4_slope = coef(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))[2],
    Night_CH4_r2 = summary(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))$r.squared
  )
lm_results <- full_join(lm_high_day,lm_high_night, join_by("Fluxing_Chamber"))

kable(lm_results, caption = "Chapada High - Flux by Chamber") %>%
  kable_styling()

```

```{r flux flag 2 - low, echo=F}
#flux slopes in table form
low <- chapada_data[["Chapada_Low_FLUX_COMB"]]
low <- low %>% filter(Flux_Status == 2, Manifold_Timer >= 180 & Manifold_Timer <= 540) 
low_day <- low %>% filter(hour(TIMESTAMP)>= 6 & hour(TIMESTAMP)<= 18)
low_night <- low %>% filter(hour(TIMESTAMP)<= 6 | hour(TIMESTAMP)>= 18)

lm_low_day <- low_day %>%
  group_by(Fluxing_Chamber) %>%
  summarise(
    Day_CO2_slope = coef(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))[2],
    Day_CO2_r2 = summary(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))$r.squared,
    Day_CH4_slope = coef(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))[2],
    Day_CH4_r2 = summary(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))$r.squared
  )

lm_low_night <- low_night %>%
  group_by(Fluxing_Chamber) %>%
  summarise(
    Night_CO2_slope = coef(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))[2],
    Night_CO2_r2 = summary(lm(as.numeric(CO2) ~ as.numeric(TIMESTAMP)))$r.squared,
    Night_CH4_slope = coef(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))[2],
    Night_CH4_r2 = summary(lm(as.numeric(CH4) ~ as.numeric(TIMESTAMP)))$r.squared
  )
lm_results <- full_join(lm_low_day,lm_low_night, join_by("Fluxing_Chamber"))

kable(lm_results, caption = "Chapada Low - Flux by Chamber") %>%
  kable_styling()

```
<br>
<br>

### Check out BME performance in each chamber:

```{r high fives for hives!, echo=F}
temp_tables <- grep("ChamberBME280", names(chapada_data), value = TRUE)
#remove ptemp
temp_data <- bind_rows(lapply(temp_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, matches("Temp\\(")) %>%
    pivot_longer(cols = -TIMESTAMP, names_to = "Sensor", values_to = "Temp") %>%
     filter(TIMESTAMP>as.POSIXct("2025-09-22 23:00:00"))%>%
    mutate(origin = table)
}))

ggplot(temp_data, aes(x = TIMESTAMP, y = as.numeric(Temp), color = Sensor)) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "BME Temperature Trends", y = "Temperature (°C)", x = "Time") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) 

rh_tables <- grep("ChamberBME280", names(chapada_data), value = TRUE)

rh_data <- bind_rows(lapply(rh_tables, function(table) {
  df <- chapada_data[[table]]
  df %>% select(TIMESTAMP, matches("RH")) %>%
    pivot_longer(cols = -TIMESTAMP, names_to = "Sensor", values_to = "RH") %>%
    mutate(origin = table)
}))

ggplot(rh_data, aes(x = TIMESTAMP, y = as.numeric(RH), color = Sensor)) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~origin) +
  labs(title = "BME Relative Humidity Trends", y = "RH (%)", x = "Time") +
  ylim(y=0, 100)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1)) 
```
