---
title: "Chapada data analysis"
author: "Abby Lewis"
date: "2026-04-14"
output: html_document
---

Kleiton- I figured it might be helpful for you to have some example code to play around with as you start working with the chapada flux data. This is very similar to the code that underlies the dashboard visualizations. Let me know if you have questions! I'm always happy to talk about code or data :)

Update 4/13/2026- I am adding a few functions from the autochamber R package I created that will help process raw flux data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)

# install.packages("remotes")
remotes::install_github("abbylewis/autochamber")
library(autochamber)
```

## Process flux files

```{r}
FILEPATHS <- c("filepath1.dat","filepath2.dat")

# Load data
data <- load_loggernet_flux_data(FILEPATHS,
                                 format = "Chapada")

# Add grouping variable and change columns
to_plot_raw <- generate_recent_raw(
  data,
  group_cols = "location",
  today = lubridate::as_date(max(data$TIMESTAMP)),
  days = 2)

# Visualize
plot_raw_data(to_plot_raw[1:500, ],
              cutoff_start = 180,
              cutoff_end = 680,
              group_cols = "location")

# Calculate flux
fluxes <- calculate_flux(data,
                         cutoff_start = 180,
                         cutoff_end = 680,
                         group_cols = "location")

# There are some funky data in here- remove weird timestamps
fluxes <- fluxes %>%
  filter(year(TIMESTAMP) > 2000)

# Add maintenance log
with_maint_log <- add_maintenance_log(
  fluxes,
  gs_url = "https://docs.google.com/spreadsheets/d/103PpjEmjLAQkov9ywjA5KxyJiIP3nEWy7V8gWVZhd1M/edit?gid=0#gid=0")

# Plot
plot_calculated_fluxes(with_maint_log,
                       group_cols = "location")
```

## CO2 and CH4

```{r load_co2_ch4}
#Load data from GitHub
latest_fluxes <- read_csv("https://raw.githubusercontent.com/abbylewis/Chapada_flux_data/refs/heads/master/processed_data/L0.csv", show_col_types = F)

#Format for plotting
fluxes_formatted <- latest_fluxes  %>%
  filter(!is.na(location),
         n>30) %>% #n is the number of data points used to calculate a flux
  mutate(TIMESTAMP = lubridate::force_tz(TIMESTAMP, tzone = "Etc/GMT-3")) %>% 
  rename(chamber = Fluxing_Chamber) %>%
  #These chamber "types" are just based on something Genevieve sent me over slack
  #They control the color of the line in the plot
  mutate(type = case_when(chamber == 90~"Termites", 
                          chamber %in% c(10,20,30) &
                            location == "low" ~ "Relatively drier",
                          chamber %in% c(10,20,30,40,50) &
                            location == "high" ~ "Relatively drier",
                          .default = "Relatively wetter"
                          ),
         chamber_long = paste0(chamber, " (", type, ")"),
         chamber = as.factor(chamber)) %>%
  # Convert to long format for plot
  pivot_longer(matches("CH4_|CO2_"), 
               names_to = c("gas", ".value"), 
               names_sep="_" ) %>%
  filter(!gas == "Flag")
```

### Plot CO2 and CH4

```{r plot_co2_ch4}
#If you plot all of the data this might be kind of overwhelming. 
# Setting start and end dates here will control what's in the plot
start_date <- as.Date("2026-01-12")
end_date <- as.Date("2026-01-14")
  
p1 <- fluxes_formatted %>%
  filter(as.Date(TIMESTAMP) <= end_date,
         as.Date(TIMESTAMP) >= start_date) %>%
  mutate(R2 = round(R2, 2)) %>%
  ggplot(aes(x = TIMESTAMP, y = slope, color = type,
             group = chamber, label = R2)) +
  geom_hline(yintercept = 0, color = "grey70") +
  geom_point(size = 0.5) +
  geom_line() +
  facet_grid(gas~location, scales = "free_y") +
  scale_x_datetime(date_labels = "%b %d") +
  theme_bw()+
  scale_color_manual(values = c("#E4D966", "#779B3E", "#673A0A"),
                     breaks = c("Relatively drier", 
                                "Relatively wetter", 
                                "Termites"),
                     name = "")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, vjust = 1.0, hjust = 1.0))

#This gives a static visualization
p1

#This gives an interactive visualization of the same data
plotly::ggplotly(p1, tooltip=c("chamber", "R2", "DateTime"))
```

## Partitioned CO2

```{r load_partitioned}
#Load data from GitHub
latest_partitioned_fluxes <- read_csv("https://raw.githubusercontent.com/abbylewis/Chapada_flux_data/refs/heads/master/processed_data/partitioned_co2.csv", show_col_types = F)

#Format for plotting
partitioned_fluxes_formatted <- latest_partitioned_fluxes %>%
  filter(!is.na(location)) %>%
  mutate(DateTime = lubridate::force_tz(DateTime, tzone = "Etc/GMT-3")) %>% 
  rename(chamber = Fluxing_Chamber) %>%
  #These chamber "types" are just based on something Genevieve sent me over slack
  #They control the color of the line in the plot
  mutate(type = case_when(chamber == 90~"Termites", 
                          chamber %in% c(10,20,30) &
                            location == "low" ~ "Relatively drier",
                          chamber %in% c(10,20,30,40,50) &
                            location == "high" ~ "Relatively drier",
                          .default = "Relatively wetter"
  ),
  chamber_long = paste0(chamber, " (", type, ")"),
  chamber = as.factor(chamber),
  GPP = ifelse(!is_day, 0, GPP)) %>%
  select(c(NEE, GPP, Reco, DateTime, chamber, location, type)) %>%
  # Convert to long format for plot
  pivot_longer(c(NEE, GPP, Reco), 
               names_to = "gas") %>%
  mutate(value = ifelse(gas == "GPP", -value, value)) %>%
  filter(!gas == "Flag")
```

### Plot partitioned CO2

```{r plot_partitioned}
#If you plot all of the data this might be kind of overwhelming. 
# Setting start and end dates here will control what's in the plot
start_date <- as.Date("2026-01-12")
end_date <- as.Date("2026-01-14")
  
p2 <- partitioned_fluxes_formatted %>%
  filter(as.Date(DateTime) <= end_date,
         as.Date(DateTime) >= start_date)%>%
  ggplot(aes(x = DateTime, y = value, color = type,
             group = chamber)) +
  geom_hline(yintercept = 0, color = "grey70") +
  geom_point(size = 0.5) +
  geom_line() +
  facet_grid(gas~location, scales = "free_y") +
  scale_x_datetime(date_labels = "%b %d") +
  theme_bw()+
  scale_color_manual(values = c("#E4D966", "#779B3E", "#673A0A"),
                     breaks = c("Relatively drier", 
                                "Relatively wetter", 
                                "Termites"),
                     name = "")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 30, vjust = 1.0, hjust = 1.0))

#This gives a static visualization
p2

#This gives an interactive visualization of the same data
plotly::ggplotly(p2, tooltip=c("chamber", "R2", "DateTime"))
```

## Drivers

```{r load_drivers}
#Load most recent data from github
bme <- read_csv("https://raw.githubusercontent.com/abbylewis/Chapada_flux_data/refs/heads/master/processed_data/bme_2025_dashboard.csv")

met <- read_csv("https://raw.githubusercontent.com/abbylewis/Chapada_flux_data/refs/heads/master/processed_data/met_2025_dashboard.csv")

redox <- read_csv("https://raw.githubusercontent.com/abbylewis/Chapada_flux_data/refs/heads/master/processed_data/redox_2025_dashboard.csv") 

teros <- read_csv("https://raw.githubusercontent.com/abbylewis/Chapada_flux_data/refs/heads/master/processed_data/teros_2025_dashboard.csv")
```

